Statistical stopping criteria for automated screening in systematic reviews

### 2022-10-19 pres/rapid-review-2022.Rmd
WWGS presentation

---
author: Max Callaghan, Finn Müller-Hansen
date: January 21, 2021
output: 
    beamer_presentation:
        latex_engine: pdflatex
        includes:
            in_header: header.tex

@@

```{r setup, include=FALSE}
set.seed(5)
knitr::opts_chunk$set(cache = F)
knitr::opts_chunk$set(global.par = TRUE)
options(max.print = 20)
xlab <- "Documents Screened" 
ylab <- "Relevant documents identified"
```
##

@@

To achieve these savings, we need to know when to stop screening: and existing approaches to deciding this are insufficient [@Shemilt2014; @Howard2020; @Jonnalagadda2013]. We need to to confident that we don't miss studies, and we need to be able to communicate this confidence.

@@

We can set up a toy dataset with 1,000 documents ($N_{tot}$), of which 100 are relevant ($\rho_{tot}$)

@@

docs <- rep(0,N_tot)
docs[1:r_tot] <- 1
docs <- sample(docs)
docs

@@

If we screen these at random, we will see 95% recall after seeing ~95% of all relevant documents

@@

It's also easy to tell when 95% recall has been achieved when we know how many relevant documents there are

\column{0.5\textwidth}
```{r echo=F}
par(pty="s")
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
abline(h=r_tot*.95)
```

@@

```{r echo=F}
par(pty="s")
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
abline(h=r_tot*.95)
```

@@

\columnsbegin
\column{0.4\textwidth}
```{r echo=F}
par(pty="s")
r_al <- cumsum(ordered_docs[1:cutoff])[cutoff]
N_al <- cutoff
K <- r_tot-r_al 
N <- N_tot - N_al
```

```{r}
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab
)
```
\column{0.6\textwidth}
$N_{AL}$ is the number of documents screened after active learning, and $\rho_{AL}$ is the number of relevant documents seen after active learning. So $N$ ($K$) is the number of (relevant) documents remaining "in the urn".
\tiny
```{r}
params <- list(
  N_al = cutoff,
  N = N_tot - N_al,
  r_al = cumsum(ordered_docs[1:cutoff])[cutoff],
  K = r_tot-r_al 
)
print(params)
```

@@

\columnsbegin
\column{0.5\textwidth}
```{r}
n <- 100
urndocs <- sample(ordered_docs[cutoff:N_tot], replace=F)
s_docs <- urndocs[1:n]
k <- cumsum(s_docs)[n]
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(s_docs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",n))
)

@@

\columnsbegin
\column{0.5\textwidth}
```{r}
vec_K_tar <- vapply(
  cumsum(urndocs)+r_al, get_ktar, numeric(1), 
  r_al=r_al, recall_target=.95
)
p <- phyper(cumsum(urndocs), vec_K_tar, N-vec_K_tar, seq(1,N))
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(urndocs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",N))
)
abline(h=r_tot*.95)
abline(v=N_al+which(p<0.05)[1])

@@

```{r}
plot(c(rep(NA,N_al),p), xlab="Documents screened")
abline(h=.05)
abline(v=N_al+which(p<0.05)[1])
```

@@

\columnsbegin
\column{0.5\textwidth}
\tiny
```{r eval=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(
  rep(1,20), col = cols, pch = 19,
  ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab=""
)
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- cumsum(tail(last_docs,i))[i]
  text(
    20.5-i*0.5,0.75+i*0.4,
    paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25)
  )
}
```
\column{0.5\textwidth}
```{r echo=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(rep(1,20), col = cols, pch = 19,ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab="")
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- cumsum(tail(last_docs,i))[i]
  text(20.5-i*0.5,0.75+i*0.4,paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25))
}
```
\columnsend

@@

@@

This presentation: \url{https://github.com/mcallaghan/rapid-screening/pres/rapid-review-esmar.Rmd}

>>>>

---
author: Max Callaghan, Finn Müller-Hansen
date: Wed 28 September 2022
output: 
    beamer_presentation:
        latex_engine: pdflatex
        includes:
            in_header: header.tex

@@

```{r setup, include=FALSE}
set.seed(5)
knitr::opts_chunk$set(cache = F)
knitr::opts_chunk$set(global.par = TRUE)
options(max.print = 35)
xlab <- "Documents Screened" 
ylab <- "Relevant documents identified"
```
##

@@

To achieve these savings, we need to know when to stop screening: and existing approaches to deciding this are insufficient [@Shemilt2014; @Howard2020; @Jonnalagadda2013]. We need to be confident that we don't miss studies, and we need to be able to communicate this confidence.

@@

We can set up a toy dataset with 2,000 documents ($N_{tot}$), of which 100 are relevant ($\rho_{tot}$)

@@

docs <- rep(0,N_tot)
docs[1:r_tot] <- 1
docs <- sample(docs, replace=F)
docs
sum(docs)

@@

If we screen these at random, we will see 95% recall after seeing ~95% of all documents

@@

@@

```{r echo=F}
par(pty="s")
cols = c(rep("black",cutoff),rep("red",N_tot-cutoff))
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified",
  col=cols
)
abline(h=r_tot*.95)
```

@@

\columnsbegin
\column{0.4\textwidth}
```{r echo=F}
par(pty="s")
r_al <- sum(ordered_docs[1:cutoff])
N_al <- cutoff
K <- r_tot-r_al 
N <- N_tot - N_al
```

```{r}
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab
)
```
\column{0.6\textwidth}
$N_{AL}$ is the number of documents screened after active learning, and $\rho_{AL}$ is the number of relevant documents seen after active learning. So $N$ ($K$) is the number of (relevant) documents remaining "in the urn".
\tiny
```{r}
params <- list(
  N_al = cutoff,
  N = N_tot - N_al,
  r_al = sum(ordered_docs[1:cutoff]),
  K = r_tot-r_al 
)
print(params)
```

@@

\columnsbegin
\column{0.5\textwidth}
```{r}
n <- 100
urndocs <- sample(ordered_docs[cutoff:N_tot], replace=F)
s_docs <- urndocs[1:n]
k <- sum(s_docs)
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(s_docs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",n))
)

@@

\columnsbegin
\column{0.5\textwidth}
```{r fig.height=6}
vec_K_tar <- vapply(
  cumsum(urndocs)+r_al, get_ktar, numeric(1), 
  r_al=r_al, recall_target=.95
)
p <- phyper(cumsum(urndocs), vec_K_tar, N-vec_K_tar, seq(1,N))
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(urndocs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",N))
)
abline(h=r_tot*.95)
abline(v=N_al+which(p<0.05)[1])

@@

```{r fig.height=6}
plot(c(rep(NA,N_al),p), xlab="Documents screened", ylab="p")
abline(h=.05)
abline(v=N_al+which(p<0.05)[1])
```

@@

\columnsbegin
\column{0.5\textwidth}
\tiny
```{r eval=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(
  rep(1,20), col = cols, pch = 19,
  ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab=""
)
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- sum(tail(last_docs,i))
  text(
    20.5-i*0.5,0.75+i*0.4,
    paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25)
  )
}
```
\column{0.5\textwidth}
```{r echo=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(rep(1,20), col = cols, pch = 19,ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab="")
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- sum(tail(last_docs,i))
  text(20.5-i*0.5,0.75+i*0.4,paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25))
}
```
\columnsend

@@

## Applications and extensions

\column{0.33\textwidth}
\only<1->{
\begin{figure}
\includegraphics[width=\linewidth]{images/doebbeling_ws.png}
\end{figure}
}

\column{0.33\textwidth}
\only<2->{
\begin{figure}
\includegraphics[width=\linewidth]{images/doebbeling_p.png}
\end{figure}
}

\column{0.33\textwidth}

\small

\begin{itemize}
  \item<1-> We have used the stopping criteria to generate massive savings (77\%) in real projects
  \item<2-> If rejecting our $H_0$ was less labour intensive we could have saved around 82\%
  \item<3-> Assuming documents are drawn at random makes our criterion over-conservative -> making more realistic assumptions (perhaps using the predictions from the models) might make it more precise.
\end{itemize}

## Open questions

\begin{itemize}
  \item<1-> Are the x\% we miss a representative sample?
  \item<2-> Can we use predictions to generate a precise estimate?
  \item<3-> Are we there yet? Can we infer the relationship between \textbf{Dataset size}, \textbf{Machine learning effectiveness}, \textbf{prevalence of relevant documents}, and \textbf{work savings}, and use this to make predictions about potential work savings?.
\end{itemize}

@@

This presentation: \url{https://github.com/mcallaghan/rapid-screening/blob/master/pres/rapid-review-esmar.Rmd}
